---
output:
  word_document: default
  html_document: default
---
### Module 3  
## Assignment 1
# Brittney Glover  

```{r}
library(tidyverse)
library(tidymodels)
library(lubridate)
```
```{r}
bike <- read_csv("bike_cleaned-2.csv")
bike = bike%>% mutate(dteday =mdy(dteday))
bike <- bike %>% mutate_if(is.character,as.factor)
bike <- bike %>% mutate(hr=as.factor(hr))
```
```{r}
set.seed(1234)
bike_split = initial_split(bike, prob = 0.70, strata = count)
train = training(bike_split)
test = testing(bike_split)
```

**How many rows of data are in each set (training and testing)?**  
There are 13036 observatios in the training data and 4343 observations in the test data.  

```{r}
bike_recipe = recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, train)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, train)
```
```{r}
summary(lm_fit$fit$fit$fit)
```

The quality of this model is fine. There are plenty predictors with significant values and the adjusted r-squared is 0.6229 which is reasonable.  

```{r}
predict_train <- predict(lm_fit, train)

ggplot(predict_train, aes(x= .pred))+
  geom_histogram()
```

There is a somewhat normal distribution of the predictions. A large range falls between 0 and 300.  

```{r}
lm_fit %>% predict(test) %>% bind_cols(test) %>% metrics(truth = count, estimate = .pred)
```
The r-squared value of the testing set is 0.6229 which is identical to the r-squared value of the training set. I would say that this is a well created model and does a good job maintaining the model with future values added. 
















