---
output:
  word_document: default
  html_document: default
---
### Assignment 3 Module 4

```{r, include = FALSE}
library(tidymodels)
library(tidyverse)
library(gridExtra)
library(caret)
library(ranger)
library(vip)
library(skimr)
```
```{r}
drug <- read_csv("drug_data-1.csv")
```
```{r}
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity","Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive","SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis","Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh","LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)
```
```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No" 
drug[drug == "CL2"] = "Yes" 
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```
```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44","45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =c("Under16", "At16", "At17", "At18", "SomeCollege","ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country,labels = c("USA", "NewZealand", "Other", "Australia","Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,labels = c("Black", "Asian", "White", "White/Black", "Other","White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>% 
  select(-ID)
  
```
```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% 
  select(!(Semer:VSA))
names(drug_clean)
```
```{r}
#skim(drug_clean)
```

There is no missingness in this data.  

```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) 
train = training(drug_split)
test = testing(drug_split)
```

## Visualizations
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle=90, vjust = 0.5))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle=90, vjust = 0.5))
grid.arrange(p1,p2,p3,p4)
```

In these plots we can see there are strong relationships with age and gender related to nicotine use. Additionally there are relationships we can see between education and country although they are ot as strong as age and gender.  

```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(p1)
```

In this plot, we can see there are strong relationships between ethnicity and nicotine use. 

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
grid.arrange(p1,p2,p3, p4, ncol = 2)
```

We can see in these plots that the Nscore, Oscore and Ascore seem to be good predictors.  

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
grid.arrange(p1,p2,p3, ncol = 2)
```

We can see that there is a strong relationship between all three of these predictors and the use of nicotine.  

```{r}
set.seed(123)
drug_folds = vfold_cv(train, v = 5)
```
```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = drug_folds,
  grid = rf_grid 
)

```
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```
```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
```{r}
final_rf_fit = fit(final_rf, train)
```
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

Open to experience, sensation seeing, and country have the highest importance in predicting nicotine use.  

```{r}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)
```
```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```
```{r}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

We see that this model performs about 9% better on the training set versus the testing set. However, both out perform the naive score.  

























