---
output:
  word_document: default
  html_document: default
---
### Module 6 Assignment 1

```{r}
library(tidyverse)
library(tidymodels)
```
```{r}
trucks <- read_csv("trucks.csv")
str(trucks)
summary(trucks)
```
```{r}
ggplot(trucks, aes(x= Distance, y=Speeding)) +
  geom_point() 
```

**Task 1**  
There does seem to be natural clustering in this data set among drivers.  

```{r}
kmeans_recipe = recipe(~ Distance + Speeding , trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks)

trucks_cleaned = bake(trucks_dummy, trucks)
```

```{r}
summary(trucks_cleaned)
summary(trucks)
```
```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 2)
```
```{r}
trucks = augment(clusters, trucks)
str(trucks)
```
```{r}
ggplot(trucks, aes(x=Distance ,y=Speeding, color=factor(.cluster))) + geom_point() 
```

**Task 3**  
There are two clusters; drivers who drive less than 100 miles a day and driver who drive more. While both clusters have a large portion that speed less than 25% of the time they are on the road, almost all drivers who drive less than 100 miles per day cap out their speeding percentage at about 60%. Where those who drive more than 100 miles per day have a decent amount of drivers who surpass the 60% mark and some who even hit 100%.  

```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>% 
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```
```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```
```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

**Task 4**  
4 clusters seems to cluster the data nicely and create clear categories for how much speeding a driver does based on the amount of miles they travel in a day.  

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

**Task 5**  
From the plot above, we can see that 4 seems to be the optimal number of clusters for this set of data.  

```{r}
set.seed(64)
clusters = kmeans(trucks_cleaned, 4)
```
```{r}
trucks = augment(clusters, trucks)
str(trucks)
```
```{r}
ggplot(trucks, aes(x=Distance ,y=Speeding, color=factor(.cluster))) + geom_point() 
```

**Task 6**  
We see in this new plot with the optimal clusters of 4, that drivers that drive less than 100 miles in a day are nicely split between those who speed less than around 15%  and those who speed more than 15% of the time that they are on the road. With the drivers who drive more than 100 miles, it is split between those who speed less than around 35% and those who speed more than 35% of the time they are on the road. The clusters are very clearly divided and the division makes sense with the data.  































